Sparse Function Space Representation of Neural Networks for Exploration and Retention

### 2023-05-10 paper/main.tex
minor

% Packages for bold math
\usepackage{bm}
\newcommand{\mathbold}[1]{\bm{#1}}
\newcommand{\mbf}[1]{\mathbf{#1}}

% Math Macros
\newcommand{\MC}{\mbf{C}}
\newcommand{\MZ}{\mbf{Z}}
\newcommand{\MV}{\mbf{V}}
\newcommand{\MX}{\mbf{X}}
\newcommand{\MA}{\mbf{A}}
\newcommand{\MK}{\mbf{K}}
\newcommand{\MI}{\mbf{I}}
\newcommand{\MH}{\mbf{H}}
\newcommand{\T}{\top}
\newcommand{\vzeros}{\mbf{0}}
\newcommand{\vtheta}[0]{\mathbold{\theta}}
\newcommand{\valpha}[0]{\mathbold{\alpha}}
\newcommand{\vkappa}[0]{\mathbold{\kappa}}
\newcommand{\vbeta}[0]{\mathbold{\beta}}
\newcommand{\vlambda}[0]{\mathbold{\lambda}}

\newcommand{\vm}{\mbf{m}}
%\newcommand{\vz}{\mbf{z}}
%\newcommand{\vf}{\mbf{f}}
\newcommand{\vu}{\mbf{u}}
\newcommand{\vx}{\mbf{x}}
\newcommand{\vy}{\mbf{y}}

\newcommand{\GP}{\mathcal{GP}}
\newcommand{\KL}[2]{\mathrm{D}_\textrm{KL} \dbar*{#1}{#2}}
\newcommand{\MKzz}{\mbf{K}_{\mbf{z}\mbf{z}}}
\newcommand{\MKxx}{\mbf{K}_{\mbf{x}\mbf{x}}}
\newcommand{\MKzx}{\mbf{K}_{\mbf{z}\mbf{x}}}
\newcommand{\MKxz}{\mbf{K}_{\mbf{x}\mbf{z}}}
\newcommand{\vkzi}{\mbf{k}_{\mbf{z}i}}
\newcommand{\vkzs}{\mbf{k}_{\mbf{z}i}}
\newcommand{\vk}{\mbf{k}}
\newcommand{\MLambda}[0]{\mathbold{\Lambda}}
\definecolor{matplotlib-blue}{HTML}{1f77b4}
\newcommand{\N}{\mathrm{N}}
\newcommand{\myexpect}{\mathbb{E}}

@@

\section{Methods} \label{sec:methods}

\textbf{Neural Network in weight space}

\textbf{Laplace approximation}

\textbf{Dual GP}

The dual parameterization states that we can decompose the optimal sparse Gaussian process posterior as follows:
\begin{align}
   \myexpect_{q_{\vu}(f)}[f_i] &= \vkzs^{\T} \valpha_{\vu} , \\
   \textrm{Var}_{q_{\vu}(f)}[f_i]  &= \kappa_{ii} - \vkzs^\top [\MKzz^{-1} - \MKzz + \vbeta_{\vu} ]^{-1} \vkzs.
\end{align}
The dual parameters consist of an $m$-length vector $\valpha_\vu$ and $m\times m$-size matrix $\vbeta_\vu$, defined as
\begin{equation} 
\begin{split}
   \vlambda & = [ \MKzz + \vbeta_{\vu}]\valpha_{\vu}  = \sum_{i \in \mathcal{D}_{\textrm{old}} \cup \mathcal{D}_{\textrm{new}}}  \vkzi \, \hat{\lambda}_{i}, \\
   \vbeta_{\vu} &= \sum_{i \in \mathcal{D}_{\textrm{old}} \cup \mathcal{D}_{\textrm{new}}} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,
\label{eq:dual_sparse}
\end{split}
\end{equation}
where $\tilde{\lambda}_i = \alpha_i - \myexpect_{q_{\vu}(f)}[f_i] \beta_i$ with $(\hat{\alpha}_i, \hat{\beta}_i)$ defined as follows,
\begin{equation}
\begin{aligned}
   \hat{\alpha}_i &= \myexpect_{q_{\vu}(f_i)} [\nabla_{f_i}\log p(y_i \mid f_i)],\\
   \hat{\beta}_i &= \myexpect_{q_{\vu}(f_i)} [ -\nabla^2_{f_i}\log p(y_i \mid f_i)],
\end{aligned} \label{eq:svgp_dual}
\end{equation}

% We start by obtaining a posterior over the weights of a trained neural network via the Laplace approximation.
% We then show how we can obtain a function-space posterior, (i.e. a GP posterior)
% by linearising the neural network around the weights Maximum a Posteriori (MAP) estimate.
% We then show that by linearising the neural network we can obtain a function-space posterior, i.e. a GP posterior.
% We paraterise our single-step dynamic model $\transitionFnWithParams : \inputDomain \rightarrow \outputDomain$
% as an $L\text{-layer}$ NN with weights $\transitionParams$.

%A Gaussian process (GP) is a distribution over real-valued functions $f(\cdot): \mathcal{X} \rightarrow \R$ defined over $\mathcal{X}$.
%Formally, a GP is defined as,
%\begin{align}
%\label{eq-gp-prior}
%  f(\cdot) \sim \mathcal{N}\left( \mu(\cdot), k(\cdot,\cdot') \right) \quad
%  \mu(\cdot): \mathcal{X} \rightarrow \R \quad
%  k(\cdot,\cdot'): \mathcal{X} \times \mathcal{X} \rightarrow \R
%\end{align}
%where $\mu(\cdot)$ denotes a mean function and $k(\cdot,\cdot')$ a positive definite covariance function, also known as a kernel.
%Given a data set $\dataset = \{\mathbf{x}_{n} \in \inputDomain, \mathbf{y}_{n} \in \outputDomain\}_{n=0}^{N}$,
%we can obtain the GP posterior via multivariate normal conditioning,
%\begin{align}
%\label{eq-gp-predictive-posterior}
%  p(f(\mathbf{x}_{*}) \mid \mathbf{y}) &= \mathcal{N}
%  % \left( f(\mathbf{x}_{*}) \mid \mathbf{A} \mathbf{m}^{*}, \mathbf{A}\mathbf{K}_{\mathbf{x}\mathbf{x}}^{-1} \mathbf{A}^{T} \right) \\
%  \left( f(\mathbf{x}_{*}) \mid \mu(\mathbf{x}_{*}) + \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}} (\mathbf{y} - \bm\mu_{\mathbf{X}})),
%  k_{**} - \mathbf{k}_{*\mathbf{x}} \left(\mathbf{K}_{\mathbf{x}\mathbf{x}} \right)^{-1} \mathbf{k}_{*\mathbf{x}}^{T} \right)
%\end{align}
%% Our GP posterior is then given by,
%\todo{add mean func to GP posterior if we're using one}
%\begin{align}
%\label{eq-gp-predictive-posterior}
%  p(f(\mathbf{x}_{*}) \mid \mathbf{y}) &= \mathcal{N}
%  % \left( f(\mathbf{x}_{*}) \mid \mathbf{A} \mathbf{m}^{*}, \mathbf{A}\mathbf{K}_{\mathbf{x}\mathbf{x}}^{-1} \mathbf{A}^{T} \right) \\
%  \left( f(\mathbf{x}_{*}) \mid \mu(\mathbf{x}_{*}) + \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}} \mu_{\mathbf{x}},
%  k_{**} - \mathbf{k}_{*\mathbf{x}} \left(\mathbf{K}_{\mathbf{x}\mathbf{x}} + \bm\Lambda(\mathbf{y} ; \mathbf{f})^{-1} \right)^{-1} \mathbf{k}_{*\mathbf{x}}^{T} \right)
%\end{align}
%where $\bm\Lambda(\mathbf{y} ; \mathbf{f}) = \nabla^{2}_{\mathbf{f} \mathbf{f}} \log p(\mathbf{y} \mid \mathbf{f})$ can be viewed as per-input noise. \todo{I'm unsure about this per input noise bit. Check with paul.}
%\todo{what about regression/classification likelhoods}
%This GP predictive posterior is computationally expensive as it requires inverting an $N\times N$ matrix.
%
%\textbf{Sparse Gaussian processes}
%% The GP predictive posterior in \cref{eq-gp-predictive-posterior} is computationally expensive as it requires inverting an $N\times N$ matrix.
%A practical and computationally appealing alternative is to augment the joint probability space with pseudo inputs $\mathbf{Z} \in \R^{M \times D}$
%and corresponding outputs
%$\mathbf{u} = f(\mathbf{Z}; \transitionParams) \in \R^{M \times C}$, known as inducing variables, where $M \ll N$.
%\todo{cite sparse gp papers}
%Instead of collapsing the inducing variables like \cite{titsiasVariational2009}, they can be treated variationally
%$p(\mathbf{u} \mid \mathbf{y}) \approx q(\mathbf{u}) = \mathcal{N}\left( \mathbf{u} \mid \mathbf{m}, \mathbf{V} \right)$, where $\mathbf{m}$ and $\mathbf{V}$ are variational parameters
%which need to be optimised \citep{hensmanGaussian2013}.
%The sparse variational GP (SVGP) predictive posterior is then given by,
%\begin{align}
%\label{eq-dual-svgp-predictive-posterior}
%  p(f(\mathbf{x}_{*}) \mid \mathbf{y})
%&\approx \mathcal{N} \left( f(\mathbf{x}_{*}) \mid \mathbf{k}_{*\mathbf{z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{m}^{*},
%  k_{**} - \mathbf{k}_{*\mathbf{z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{k}_{*\mathbf{z}}^{T}
%  % \mathbf{k}_{*\inducingInput} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{k}_{*\inducingInput}^{T}
%  + \mathbf{k}_{*\mathbf{z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}}  \mathbf{V}^{*}  \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{k}_{*\mathbf{z}}^{T}
%  \right)
%\coloneqq q_{\inducingVariable}(f(\mathbf{x}_{*}))
%\end{align}
%
%
%\textbf{Deep learning}
%Given the data set $\dataset = \{\mathbf{x}_{n} \in \inputDomain, \mathbf{y}_{n} \in \outputDomain\}_{n=0}^{N}$,
%the goal of (supervised) deep learning, is to train the weights $\transitionParams \in \R^{P}$ of an $L\text{-layer}$ NN
%$f : \inputDomain \rightarrow \outputDomain$ to minimize the (regularized) empirical risk,
%\begin{align} \label{eq-empirical-risk}
%  \transitionParams^{*} = \arg \min_{\transitionParams \in \R^{D}} \mathcal{L}(\dataset;\transitionParams) = \arg \min_{\transitionParams \in \R^{D}} \left(
%  \sum_{n=0}^{N-1} l(f(\mathbf{x}_{n} ; \transitionParams), \mathbf{y}_{n}) + R(\transitionParams)  \right).
%\end{align}
%In practice, it is common to use the mean squared error (MSE) loss
%$l(f(\mathbf{x}_{n} ; \transitionParams),\mathbf{y}_{n}) = \|f(\mathbf{x}_{n} ; \transitionParams) - \mathbf{y}_{n} \|^{2}_{2}$
%and to use the weight decay regularizer $R(\transitionParams)=\frac{1}{2}\gamma^{-2}\|\transitionParams\|^{2}_{2}$.
%However, the practioner is free to choose any loss and regularizer. \todo{I think?}
%% \textbf{Bayesian neural networks}
%% \textbf{Weight space to function space}
%From a Bayesian perspective, we can interpret the terms in \cref{eq-empirical-risk} as the log-prior and the i.i.d. log likelihood, i.e.,
%\begin{align} \label{eq-log-prior}
%  R(\transitionParams) &= \underbrace{\log p(\transitionParams)}_{\text{log prior}} = \log \mathcal{N}(\transitionParams \mid 0, \gamma^{2} \mathbf{I}) \\
%  l(f(\mathbf{x}_{n} ; \transitionParams), \mathbf{y}_{n}) &= \underbrace{\log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}; \transitionParams))}_{\text{log likelihood}}
%  = \log \mathcal{N} \left( \mathbf{y}_{n} \mid f(\mathbf{x}_{n}; \transitionParams), \mathbf{I} \right) \label{eq-log-likelihood}
%\end{align}
%
%\textbf{NN to GP}
%Given this interpretation, we can represent the weight space prior from \cref{eq-log-prior}, in
%function space, by linearising the BNN around $\transitionParams^{*}$ and interpretting it as a GP.
%The GP prior is then given by,
%\begin{align} \label{eq-laplace-approx-function-space}
%  % \transitionFn_{\transitionParams_{i}}(\cdot) \sim \mathcal{N} \left( \mu_{\transitionParams_{i}^{*}}(\cdot), K_{\transitionParams_{i}^{*}}(\cdot, \cdot') \right) \qquad
%  f(\cdot ;\transitionParams) \sim \mathcal{N} \left( \mu(\cdot), k(\cdot, \cdot') \right) \quad
%  % \mu_{\transitionParams_{i}^{*}}(\cdot)
%  \mu(\cdot)
%  = 0 \quad
%  % = f(\cdot ;\transitionParams^{*}) \quad
%  % K_{\transitionParams_{i}^{*}}(\cdot, \cdot')
%  k(\cdot, \cdot')
%  = \frac{1}{\gamma^{2}} \mathbf{J}(\cdot) \mathbf{J}(\cdot')^{T},
%\end{align}
%where the kernel is the neural tangent kernel (NTK) \citep{immerImprovingPredictionsBayesian2021},
%\todo{what mean function shoudl this use???}
%\todo{NTK times some factor??}
%with Jacobian given by $\mathbf{J}(\cdot) = \odv{\transitionFn(\cdot; \transitionParams)}{\transitionParams}_{\transitionParams=\transitionParams^{*}}$.
%% The kernel in \cref{eq-laplace-approx-function-space} is the neural tangent kernel (NTK)
%% $k(\mathbf{x}, \mathbf{x}') = \sigma_{0}^{2} J_{\transitionParams^{*}}(\mathbf{x}) J_{\transitionParams^{*}}(\mathbf{x}')^{T}$
%% \citep{immerImprovingPredictionsBayesian2021}
%% to formulate the GP posterior $p(f(\mathbf{X} ;\transitionParams_{0:i}) \mid \dataset_{0:i})$.
%Given this GP prior, we can use \cref{eq-gp-predictive-posterior} to obtain our GP posterior by conditioning on the data.
%% Given our GP prior, we can use the properties of multivariate normals to obtain our GP posterior by conditioning on the data.
%% where $\mathbf{A} = \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}}$
%In contrast to conventional GP kernels, the NTK does not have any hyperparameters which need to be learned.
%Intuitively, we can view the optimisation in \cref{eq-empirical-risk} as learning our kernel.
%As a result, our NTK may be highly non stationary as it is dependent on the NN architecture, for example, the activation functions.
%It is worth noting that the NTK linearises the network around the optimised parameters $\transitionParams^{*}$,
%so the function space prior (and thus posterior) is only a locally linear approximation.
%
%We now have a method to obtain a GP posterior from a trained DNN.
%% In contrast to other approaches, our formulation has not restricted us to using a weight-space posterior obtained from the Laplace approximation.
%In contrast to other approaches, our formulation has not restricted us to using a weight-space posterior from a BNN.
%However, in practice, the GP posterior is not very useful as it requires inverting an $N\times N$ matrix which has complexity $\mathcal{O}(N^{3})$.
%The SVGP in \cref{eq-dual-svgp-predictive-posterior} offers a solution to this problem but it is not immediately clear how to set $\mathbf{m}$ and $\mathbf{V}$.
%In the next section, we show how we can directly obtain an SVGP posterior (i.e. set $\mathbf{m}$ and $\mathbf{V}$) directly from both
%1) a trained NN and 2) a BNN weight-space posterior.
%% $\mathbf{m}$ and $\mathbf{V}$
%% However, it is not obvious how we can use a SVGP to set $\mathbf{m}$ and $\mathbf{V}$
%
%
%
%% In contrast, to previous work, this approach did not require us Laplace approximation. we will not restrict our
%% It is worth noting that the NTK linearises the network around the optimised parameters $\transitionParams^{*}$,
%% so the function space prior (and thus posterior) is only a locally linear approximation.
%% In contrast to conventional GP kernels, the NTK does not have any hyperparameters which need to be learned.
%% It is also worth noting that the NTK may be highly non stationary and is dependent on the NN architecture, for example, the activation functions.
%
%
%\section{Methods}
%In the next section we show how we can obtain a SVGP directly from 1) a trained NN and 2) a BNN posterior.
%
%We follow \cite{csatoSparseOnlineGaussian2002} and express our GP posterior in the dual parameter space.
%However, we consider the variational formulation introduced by \cite{adamDualParameterizationSparse2021}, where the optimal variational parameters are given by,
%\begin{align} \label{eq-dual-params}
%\mathbf{m}^{*} = \mathbf{V}^{*}\bm\alpha^{*} \quad \mathbf{V}^{*} = [\mathbf{K}_{\mathbf{z}\mathbf{z}}^{-1} + \bm\beta^{*}]
%\end{align}
%Importantly, we can calculate the natural parameters $(\bm\alpha, \bm\beta)$ in closed form,
%\todo{paul needs to update \cref{eq-dual-svgp-params}}
%\begin{align} \label{eq-dual-svgp-params}
%\bm\alpha_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right] \\
%\bm\beta_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right].
%\end{align}
%This simplifies SVGP inference \citep{hensmanGaussian2013} as it no longer requires the variational parameters $(\mathbf{m}, \mathbf{V})$ to be optimised.
%
%% Importantly, \cite{changFantasizingDualGPs2022} show that in the dual space,
%% conditioning on new observations reduces to suming the dual variables from the previous time
%% step with an update, given by,
%% \todo{pick best way to show new data (using superscript new or just time indexing?)}
%% \begin{align} \label{eq-dual-update-svgp}
%% \bm\alpha_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right] \\

>>>>

@@

\section{Background} \label{sec:background}
Sequential decision making requires uncertainty estimates for effective exploration.

In this section, we recap how can obtain a GP posterior from a trained deep neural network (DNN).
% We start by obtaining a posterior over the weights of a trained neural network via the Laplace approximation.
% We then show how we can obtain a function-space posterior, (i.e. a GP posterior)
% by linearising the neural network around the weights Maximum a Posteriori (MAP) estimate.
% We then show that by linearising the neural network we can obtain a function-space posterior, i.e. a GP posterior.
% We paraterise our single-step dynamic model $\transitionFnWithParams : \inputDomain \rightarrow \outputDomain$
% as an $L\text{-layer}$ NN with weights $\transitionParams$.
\textbf{Gaussian processes}
A Gaussian process (GP) is a distribution over real-valued functions $f(\cdot): \mathcal{X} \rightarrow \R$ defined over $\mathcal{X}$.
Formally, a GP is defined as,
\begin{align}
\label{eq-gp-prior}
  f(\cdot) \sim \mathcal{N}\left( \mu(\cdot), k(\cdot,\cdot') \right) \quad
  \mu(\cdot): \mathcal{X} \rightarrow \R \quad
  k(\cdot,\cdot'): \mathcal{X} \times \mathcal{X} \rightarrow \R
\end{align}
where $\mu(\cdot)$ denotes a mean function and $k(\cdot,\cdot')$ a positive definite covariance function, also known as a kernel.
Given a data set $\dataset = \{\mathbf{x}_{n} \in \inputDomain, \mathbf{y}_{n} \in \outputDomain\}_{n=0}^{N}$,
we can obtain the GP posterior via multivariate normal conditioning,
\begin{align}
\label{eq-gp-predictive-posterior}
  p(f(\mathbf{x}_{*}) \mid \mathbf{y}) &= \mathcal{N}
  % \left( f(\mathbf{x}_{*}) \mid \mathbf{A} \mathbf{m}^{*}, \mathbf{A}\mathbf{K}_{\mathbf{x}\mathbf{x}}^{-1} \mathbf{A}^{T} \right) \\
  \left( f(\mathbf{x}_{*}) \mid \mu(\mathbf{x}_{*}) + \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}} (\mathbf{y} - \bm\mu_{\mathbf{X}})),
  k_{**} - \mathbf{k}_{*\mathbf{x}} \left(\mathbf{K}_{\mathbf{x}\mathbf{x}} \right)^{-1} \mathbf{k}_{*\mathbf{x}}^{T} \right)
\end{align}
% Our GP posterior is then given by,
\todo{add mean func to GP posterior if we're using one}
\begin{align}
\label{eq-gp-predictive-posterior}
  p(f(\mathbf{x}_{*}) \mid \mathbf{y}) &= \mathcal{N}
  % \left( f(\mathbf{x}_{*}) \mid \mathbf{A} \mathbf{m}^{*}, \mathbf{A}\mathbf{K}_{\mathbf{x}\mathbf{x}}^{-1} \mathbf{A}^{T} \right) \\
  \left( f(\mathbf{x}_{*}) \mid \mu(\mathbf{x}_{*}) + \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}} \mu_{\mathbf{x}},
  k_{**} - \mathbf{k}_{*\mathbf{x}} \left(\mathbf{K}_{\mathbf{x}\mathbf{x}} + \bm\Lambda(\mathbf{y} ; \mathbf{f})^{-1} \right)^{-1} \mathbf{k}_{*\mathbf{x}}^{T} \right)
\end{align}
where $\bm\Lambda(\mathbf{y} ; \mathbf{f}) = \nabla^{2}_{\mathbf{f} \mathbf{f}} \log p(\mathbf{y} \mid \mathbf{f})$ can be viewed as per-input noise. \todo{I'm unsure about this per input noise bit. Check with paul.}
\todo{what about regression/classification likelhoods}
This GP predictive posterior is computationally expensive as it requires inverting an $N\times N$ matrix.

\textbf{Sparse Gaussian processes}
% The GP predictive posterior in \cref{eq-gp-predictive-posterior} is computationally expensive as it requires inverting an $N\times N$ matrix.
A practical and computationally appealing alternative is to augment the joint probability space with pseudo inputs $\mathbf{Z} \in \R^{M \times D}$
and corresponding outputs
$\mathbf{u} = f(\mathbf{Z}; \transitionParams) \in \R^{M \times C}$, known as inducing variables, where $M \ll N$.
\todo{cite sparse gp papers}
Instead of collapsing the inducing variables like \cite{titsiasVariational2009}, they can be treated variationally
$p(\mathbf{u} \mid \mathbf{y}) \approx q(\mathbf{u}) = \mathcal{N}\left( \mathbf{u} \mid \mathbf{m}, \mathbf{V} \right)$, where $\mathbf{m}$ and $\mathbf{V}$ are variational parameters
which need to be optimised \citep{hensmanGaussian2013}.
The sparse variational GP (SVGP) predictive posterior is then given by,
\begin{align}
\label{eq-dual-svgp-predictive-posterior}
  p(f(\mathbf{x}_{*}) \mid \mathbf{y})
&\approx \mathcal{N} \left( f(\mathbf{x}_{*}) \mid \mathbf{k}_{*\mathbf{z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{m}^{*},
  k_{**} - \mathbf{k}_{*\mathbf{z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{k}_{*\mathbf{z}}^{T}
  % \mathbf{k}_{*\inducingInput} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{k}_{*\inducingInput}^{T}
  + \mathbf{k}_{*\mathbf{z}} \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}}  \mathbf{V}^{*}  \mathbf{K}^{-1}_{\mathbf{z}\mathbf{z}} \mathbf{k}_{*\mathbf{z}}^{T}
  \right)
\coloneqq q_{\inducingVariable}(f(\mathbf{x}_{*}))
\end{align}

\textbf{Deep learning}
Given the data set $\dataset = \{\mathbf{x}_{n} \in \inputDomain, \mathbf{y}_{n} \in \outputDomain\}_{n=0}^{N}$,
the goal of (supervised) deep learning, is to train the weights $\transitionParams \in \R^{P}$ of an $L\text{-layer}$ NN
$f : \inputDomain \rightarrow \outputDomain$ to minimize the (regularized) empirical risk,
\begin{align} \label{eq-empirical-risk}
  \transitionParams^{*} = \arg \min_{\transitionParams \in \R^{D}} \mathcal{L}(\dataset;\transitionParams) = \arg \min_{\transitionParams \in \R^{D}} \left(
  \sum_{n=0}^{N-1} l(f(\mathbf{x}_{n} ; \transitionParams), \mathbf{y}_{n}) + R(\transitionParams)  \right).
\end{align}
In practice, it is common to use the mean squared error (MSE) loss
$l(f(\mathbf{x}_{n} ; \transitionParams),\mathbf{y}_{n}) = \|f(\mathbf{x}_{n} ; \transitionParams) - \mathbf{y}_{n} \|^{2}_{2}$
and to use the weight decay regularizer $R(\transitionParams)=\frac{1}{2}\gamma^{-2}\|\transitionParams\|^{2}_{2}$.
However, the practioner is free to choose any loss and regularizer. \todo{I think?}
% \textbf{Bayesian neural networks}
% \textbf{Weight space to function space}
From a Bayesian perspective, we can interpret the terms in \cref{eq-empirical-risk} as the log-prior and the i.i.d. log likelihood, i.e.,
\begin{align} \label{eq-log-prior}
  R(\transitionParams) &= \underbrace{\log p(\transitionParams)}_{\text{log prior}} = \log \mathcal{N}(\transitionParams \mid 0, \gamma^{2} \mathbf{I}) \\
  l(f(\mathbf{x}_{n} ; \transitionParams), \mathbf{y}_{n}) &= \underbrace{\log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}; \transitionParams))}_{\text{log likelihood}}
  = \log \mathcal{N} \left( \mathbf{y}_{n} \mid f(\mathbf{x}_{n}; \transitionParams), \mathbf{I} \right) \label{eq-log-likelihood}
\end{align}

\textbf{NN to GP}
Given this interpretation, we can represent the weight space prior from \cref{eq-log-prior}, in
function space, by linearising the BNN around $\transitionParams^{*}$ and interpretting it as a GP.
The GP prior is then given by,
\begin{align} \label{eq-laplace-approx-function-space}
  % \transitionFn_{\transitionParams_{i}}(\cdot) \sim \mathcal{N} \left( \mu_{\transitionParams_{i}^{*}}(\cdot), K_{\transitionParams_{i}^{*}}(\cdot, \cdot') \right) \qquad
  f(\cdot ;\transitionParams) \sim \mathcal{N} \left( \mu(\cdot), k(\cdot, \cdot') \right) \quad
  % \mu_{\transitionParams_{i}^{*}}(\cdot)
  \mu(\cdot)
  = 0 \quad
  % = f(\cdot ;\transitionParams^{*}) \quad
  % K_{\transitionParams_{i}^{*}}(\cdot, \cdot')
  k(\cdot, \cdot')
  = \frac{1}{\gamma^{2}} \mathbf{J}(\cdot) \mathbf{J}(\cdot')^{T},
\end{align}
where the kernel is the neural tangent kernel (NTK) \citep{immerImprovingPredictionsBayesian2021},
\todo{what mean function shoudl this use???}
\todo{NTK times some factor??}
with Jacobian given by $\mathbf{J}(\cdot) = \odv{\transitionFn(\cdot; \transitionParams)}{\transitionParams}_{\transitionParams=\transitionParams^{*}}$.
% The kernel in \cref{eq-laplace-approx-function-space} is the neural tangent kernel (NTK)
% $k(\mathbf{x}, \mathbf{x}') = \sigma_{0}^{2} J_{\transitionParams^{*}}(\mathbf{x}) J_{\transitionParams^{*}}(\mathbf{x}')^{T}$
% \citep{immerImprovingPredictionsBayesian2021}
% to formulate the GP posterior $p(f(\mathbf{X} ;\transitionParams_{0:i}) \mid \dataset_{0:i})$.
Given this GP prior, we can use \cref{eq-gp-predictive-posterior} to obtain our GP posterior by conditioning on the data.
% Given our GP prior, we can use the properties of multivariate normals to obtain our GP posterior by conditioning on the data.
% where $\mathbf{A} = \mathbf{k}_{*\mathbf{x}} \mathbf{K}^{-1}_{\mathbf{x}\mathbf{x}}$
In contrast to conventional GP kernels, the NTK does not have any hyperparameters which need to be learned.
Intuitively, we can view the optimisation in \cref{eq-empirical-risk} as learning our kernel.
As a result, our NTK may be highly non stationary as it is dependent on the NN architecture, for example, the activation functions.
It is worth noting that the NTK linearises the network around the optimised parameters $\transitionParams^{*}$,
so the function space prior (and thus posterior) is only a locally linear approximation.

We now have a method to obtain a GP posterior from a trained DNN.
% In contrast to other approaches, our formulation has not restricted us to using a weight-space posterior obtained from the Laplace approximation.
In contrast to other approaches, our formulation has not restricted us to using a weight-space posterior from a BNN.
However, in practice, the GP posterior is not very useful as it requires inverting an $N\times N$ matrix which has complexity $\mathcal{O}(N^{3})$.
The SVGP in \cref{eq-dual-svgp-predictive-posterior} offers a solution to this problem but it is not immediately clear how to set $\mathbf{m}$ and $\mathbf{V}$.
In the next section, we show how we can directly obtain an SVGP posterior (i.e. set $\mathbf{m}$ and $\mathbf{V}$) directly from both
1) a trained NN and 2) a BNN weight-space posterior.
% $\mathbf{m}$ and $\mathbf{V}$
% However, it is not obvious how we can use a SVGP to set $\mathbf{m}$ and $\mathbf{V}$

% In contrast, to previous work, this approach did not require us Laplace approximation. we will not restrict our
% It is worth noting that the NTK linearises the network around the optimised parameters $\transitionParams^{*}$,
% so the function space prior (and thus posterior) is only a locally linear approximation.
% In contrast to conventional GP kernels, the NTK does not have any hyperparameters which need to be learned.
% It is also worth noting that the NTK may be highly non stationary and is dependent on the NN architecture, for example, the activation functions.

\section{Methods}
In the next section we show how we can obtain a SVGP directly from 1) a trained NN and 2) a BNN posterior.

We follow \cite{csatoSparseOnlineGaussian2002} and express our GP posterior in the dual parameter space.
However, we consider the variational formulation introduced by \cite{adamDualParameterizationSparse2021}, where the optimal variational parameters are given by,
\begin{align} \label{eq-dual-params}
\mathbf{m}^{*} = \mathbf{V}^{*}\bm\alpha^{*} \quad \mathbf{V}^{*} = [\mathbf{K}_{\mathbf{z}\mathbf{z}}^{-1} + \bm\beta^{*}]
\end{align}
Importantly, we can calculate the natural parameters $(\bm\alpha, \bm\beta)$ in closed form,
\todo{paul needs to update \cref{eq-dual-svgp-params}}
\begin{align} \label{eq-dual-svgp-params}
\bm\alpha_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right] \\
\bm\beta_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right].
\end{align}
This simplifies SVGP inference \citep{hensmanGaussian2013} as it no longer requires the variational parameters $(\mathbf{m}, \mathbf{V})$ to be optimised.

% Importantly, \cite{changFantasizingDualGPs2022} show that in the dual space,
% conditioning on new observations reduces to suming the dual variables from the previous time
% step with an update, given by,
% \todo{pick best way to show new data (using superscript new or just time indexing?)}
% \begin{align} \label{eq-dual-update-svgp}
% \bm\alpha_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right] \\
% \bm\beta_{n} &=  \nabla_{\meanParam{1}} \mathbb{E}_{q_{\inducingVariable}(f(\mathbf{x}_{n}))} \left[ \log p(\mathbf{y}_{n} \mid f(\mathbf{x}_{n}) ) \right]
% \end{align}
